---
title: "Airbnb"
author: "Luke O'Connell & Yvonne Quach"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small

library(knitr)

opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = TRUE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```



```{r}
# Load required libraries
library(dplyr)
library(tidyr)
library(tidyverse)
library(randomForest)
library(caret)
library(stringr)
library(forcats)

# Load the dataset

#Yvonne, here are your load instructions: 
#LA_listings <- read_csv("LA_listings.csv", show_col_types = FALSE)

#Luke, here are your load instructions: 
LA_listings <- read_csv("/Users/leo/Desktop/BUS659/project/LA_listings.csv", show_col_types = FALSE)
```

Start: Price optimization... Predict optimal nightly rates based on listing’s characteristics, like room type, amenities, reviews, and location.

Since there are 3 objectives, all of which will require testing, training, and likely creation of similiar data sets, all data groups created for price optimization will begin with "opt_". 

```{r}

#Select relevant columns and clean them for modeling
opt_LA_price <- LA_listings %>% 
  select(price, room_type, neighbourhood_cleansed, number_of_reviews, review_scores_rating, minimum_nights, availability_365, latitude, longitude, amenities, bedrooms, bathrooms, beds) %>% 
  mutate(
    
    # Convert price string to numeric
    price = as.numeric(gsub("[$,]", "", price)),
    
    # Convert categorical variables to factors
    room_type = as.factor(room_type),
    neighbourhood_cleansed = as.factor(neighbourhood_cleansed),
    
    #Convert bedroom/bathroom/bed counts to numeric
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    
    # Create a new numeric feature: count of amenities
    amenities_count = str_count(amenities, ",") +1,
    
    #Price per bedroom (if bedroom is not 0), indicates price based on listing size
    price_per_bedroom = ifelse(bedrooms > 0, price / bedrooms, NA),
    
     #Price per bed, indicates price based on listing size
    price_per_bed = ifelse(beds > 0, price / beds, NA),
    
    #Indicator if listing is entire home, for price segmentation
    is_entire_home = ifelse(room_type == "Entire home/apt", 1, 0),
    
    #Log-transformed price to address skewness for sensitive distributions
    log_price = log1p(price)
  ) %>% 
  
  # Drop any rows that have missing values in the selected columns
  drop_na() %>% 
  
  # Drop the raw amenities text column 
  select(-amenities)
  
# Random Forest cannot handle factors with more than 53 levels, so we group smaller neighborhoods into an "Other" category
opt_top_neighborhoods <- opt_LA_price %>% 
  count(neighbourhood_cleansed, sort = TRUE) %>% 
  slice(1:10) %>% 
  pull(neighbourhood_cleansed) %>% 
  as.character()

opt_LA_price <- opt_LA_price %>% 
  mutate(neighbourhood_cleansed = fct_other(neighbourhood_cleansed, keep = opt_top_neighborhoods))
```


```{r}

# Split the data into training and testing sets

# Load rsample
library(rsample)

# Use rsample's initial_split() to create an 80/20 train/test split
set.seed(123)
opt_split <- initial_split(opt_LA_price, prop = 0.8)
    
# Extract training and testing sets
opt_LA_train <- training(opt_split)
opt_LA_test <- testing(opt_split)
 
```

```{r}
# Train the Random Forest model

# Fit a random forrest regression model to predict nightly price
opt_rf_model <- randomForest(
  
  # Predice price using all other variables
  price ~ .,
  
  # Use training data only
  data = opt_LA_train,
  
  # Number of trees in the forest
  ntree = 100,
  
  # Track feature importance for interpretation
  importance = TRUE
)

```


```{r}

# Predict on the test data
predictions <- predict(opt_rf_model, newdata = opt_LA_test)

# Calculate performance metrics
mae <- mean(abs(predictions - opt_LA_test$price))
rmse <- sqrt(mean((predictions - opt_LA_test$price)^2))
r2 <- cor(predictions, opt_LA_test$price)^2

# Print the results
cat("MAE:", round(mae, 2), "\n")
cat("RMSE:", round(rmse, 2), "\n")
cat("R2:", round(r2, 3), "\n")

```
MAE: 4.05 (much lower error)
RMSE: 83.49 (lower is better... not perfect, but better)
R2: 0.975 (explains 0.415 more of the variance)

```{r}

#The new MAE is $69.52, so on average, our predictions are off by $70 per listing, which is a slight improvement from before. The RMSE is $273.42, so our model still has some large errors on more expensive listings, but slightly better than before. The r-squared is 0.587, which means our model now explains 59% of the variance in Airbnb prices, a slight improvement from 56%. 
```

End: Price optimization

Start: Revenue forecasting... Estimate annual revenue per listing using reported attributes of each listing for hosts to make informed decisions about setting prices and availability on their properties. Estimate expected future revenue for listings using pricing, availability, and booking history.

Since there are 3 objectives, all of which will require testing, training, and likely creation of similar data sets, all data groups created for price optimization will begin with "rev_". 

```{r}

# Select relevant columns and clean them for modeling

rev_LA_avail <- LA_listings %>%
  select(price, room_type, neighbourhood_cleansed, number_of_reviews, review_scores_rating,
         minimum_nights, availability_365, latitude, longitude, amenities,
         bedrooms, bathrooms, beds) %>%
  mutate(
    # Convert price string to numeric
    price = as.numeric(gsub("[$,]", "", price)),
    
    # Create proxy for bookings, allowing number of reviews to stand in for bookings if no booking data is available
    #This assumes a correlation between reviews and bookings, which likely underestimates revenue for newer listings
    proxy_bookings = number_of_reviews,
    
    # Create revenue variable and log-transformed revenue
    revenue = price * proxy_bookings,
    log_revenue = log1p(revenue),
    
    # Isolate availability by singular days
    availability_rate = availability_365 / 365,
    
    # Convert categorical variables to factors
    room_type = as.factor(room_type),
    neighbourhood_cleansed = as.factor(neighbourhood_cleansed),
    
    # Convert bedroom/bathroom/bed counts to numeric
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    
    # Create a new numeric feature: count of amenities
    amenities_count = str_count(amenities, ",") + 1
  ) %>%
  select(-amenities) %>%  # Drop raw amenities text
  drop_na()
```

```{r}

# Create an 80/20 train/test split

rev_split <- initial_split(rev_LA_avail, prop = 0.8)

# Extract training and testing sets
rev_train <- training(rev_split)
rev_test  <- testing(rev_split)

```

```{r}

# Load additional required libraries for Elastic Net

library(glmnet)
library(glmnetUtils)

# We did not learn model.matrix in class, however, it was covered in BDS's section on regression and model selection in chapter 2

# Create revenue training and testing matrices
rev_train_matrix <- model.matrix(~ . -1, data = rev_train %>% select(-log_revenue))
rev_train_target <- rev_train$log_revenue

rev_test_matrix <- model.matrix(~ . -1, data = rev_test %>% select(-log_revenue))
rev_test_target <- rev_test$log_revenue

```

```{r}

# Fit by using cross-validation across multiple alpha values
set.seed(123)

enet_mod <- cva.glmnet(
  x = rev_train_matrix,
  y = rev_train_target,
  alpha = seq(0, 1, length = 11),
  family = "gaussian"
)

# I pulled this from regularization slide page 55. Helper function to find the alpha with lowest cross-validation error
rev_get_alpha <- function(fit) {
  alpha <- fit$alpha
  error <- sapply(fit$modlist, function(mod) {min(mod$cvm)})
  alpha[which.min(error)]
}

# Helper function to extract alpha, lambda.min, lambda.1se, and error
rev_get_model_params <- function(fit) {
  alpha <- fit$alpha
  lambdaMin <- sapply(fit$modlist, `[[`, "lambda.min")
  lambdaSE <- sapply(fit$modlist, `[[`, "lambda.1se")
  error <- sapply(fit$modlist, function(mod) {min(mod$cvm)})
  best <- which.min(error)
  data.frame(alpha = alpha[best], lambdaMin = lambdaMin[best], lambdaSE = lambdaSE[best], error = error[best])
}

# Identify best alpha and lambda values
rev_best_alpha <- rev_get_alpha(enet_mod)
print(paste("Best alpha:", rev_best_alpha))

rev_model_params <- rev_get_model_params(enet_mod)
print(rev_model_params)

# Extract best model using optimal alpha
rev_best_model <- enet_mod$modlist[[which(enet_mod$alpha == rev_best_alpha)]]

# Predict log revenue using the best model and optimal lambda
rev_enet_predictions <- predict(
  rev_best_model,
  s = rev_model_params$lambdaMin,
  newx = rev_test_matrix
)

# Evaluate model performance on the test set
rev_enet_mae <- mean(abs(rev_enet_predictions - rev_test_target))
rev_enet_rmse <- sqrt(mean((rev_enet_predictions - rev_test_target)^2))
rev_enet_r2 <- cor(rev_enet_predictions, rev_test_target)^2

# Output performance results
cat("Elastic Net (best alpha =", rev_best_alpha, ") - MAE:", round(rev_enet_mae, 2), "\n")
cat("Elastic Net (best alpha =", rev_best_alpha, ") - RMSE:", round(rev_enet_rmse, 2), "\n")
cat("Elastic Net (best alpha =", rev_best_alpha, ") - R²:", round(rev_enet_r2, 3), "\n")
```

MAE: 0.95 
RMSE: 1.16 
R2: 0.616 

Since we are unable to see actual revenue per listing, and it can be assumed that revenue is highly skewed based on seasonality and dynamic pricing, most of these numbers are fine for now. We should fine tune them. We may try adding superhost status, property_type?

We may also want to consider removing log values... If we want to see actual dollar value of revenue. Luke will mess with that later. 

End: Revenue forecasting



Start: Market segmentation

Since there are 3 objectives, all of which will require testing, training, and likely creation of similiar data sets, all data groups created for market segmentation will begin with "mktseg_". 

Yvonne- 
1. Check out these slides from Hersh's prior classes. They are slightly different from each other. Looks like this was the last component taught in class in 2019ish: https://github.com/jonhersh/MGSC310/blob/master/slides/Class%2020%20-%20Clustering.pdf, 

and 

https://github.com/jonhersh/Machine_Learning_For_Managers/blob/master/slides/Lecture%2011%20-%20Data%20Cleaning%20and%20Unsupervised%20Learning.pdf

2. His syllabus from this same semester suggests pages 385-399 in this free book: https://static1.squarespace.com/static/5ff2adbe3fe4fe33db902812/t/6009dd9fa7bc363aa822d2c7/1611259312432/ISLR+Seventh+Printing.pdf

Luke, check back in on "ggcorrplot"
```{r}

# Select relevant columns and clean them for modeling, including both features of the property and indications of demand

mktseg_cluster_data <- LA_listings %>%
  select(price, bedrooms, bathrooms, beds, review_scores_rating, availability_365, amenities) %>%
  mutate(
    # Convert price to numeric by removing formatting
    price = as.numeric(gsub("[$,]", "", price)),
    
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    review_scores_rating = as.numeric(review_scores_rating),
    availability_365 = as.numeric(availability_365),
    
    amenities_count = str_count(amenities, ",") + 1
  ) %>%
  select(-amenities) %>%

  drop_na()

```

```{r}

# Standardize features before clustering, K-Means requires variables to be on a similar scale
# Scaling so that variables measured by different units (price vs. rating) contribute equally
mktseg_cluster_data_scaled <- scale(mktseg_cluster_data)

# Elbow method to choose k (number of clusters)
# "Elbow" looks at total within-cluster sum of squares (WSS) across different k values
# The "elbow" point suggests diminishing returns from additional clusters
set.seed(123)
wss <- sapply(1:10, function(k){
  kmeans(mktseg_cluster_data_scaled, centers = k, nstart = 10)$tot.withinss
})

# Visualize the WSS to identify optimal number of clusters
plot(1:10, wss, type = "b", pch = 19,
     xlab = "Number of Clusters K",
     ylab = "Total Within Sum of Squares",
     main = "Elbow Method for Optimal K")

```

```{r}

# Apply K-means clustering. Based on the elbow plot, select a k = x (adjust x to different # if appropriate)
mktseg_kmeans_result <- kmeans(mktseg_cluster_data_scaled, centers = 4, nstart = 25)

# Add the resulting cluster assignments as a new variable,so now each row belongs to one of the k (4) market segments (labeled as 1 through 4)
# Instead of assigning directly to LA_clean (which causes a row mismatch), we assign segment labels only to the dataset used for clustering

# Track row IDs for consistent matching with clustering results
mktseg_clustered <- mktseg_cluster_data %>%
  mutate(row_id = as.integer(rownames(.))) %>%
  mutate(segment = as.factor(mktseg_kmeans_result$cluster))

# Add row IDs to the original dataset to enable safe join
mktseg_LA_listings <- LA_listings %>%
  mutate(row_id = row_number())

# Create a new dataset with segment assignments (no change to original LA_listings)
mktseg_LA_segmented <- mktseg_LA_listings %>%
  left_join(select(mktseg_clustered, row_id, segment, amenities_count), by = "row_id") %>%
  select(-row_id)

```

```{r}

# Summarize the characteristics of each segment, helps interpret how different segments differ in key features (e.g., price, size, amenities)
mktseg_segment_summary <- mktseg_LA_segmented %>%
  group_by(segment) %>%
  summarise(
    avg_price = mean(price),
    avg_bedrooms = mean(bedrooms),
    avg_bathrooms = mean(bathrooms),
    avg_beds = mean(beds),
    avg_rating = mean(review_scores_rating, na.rm = TRUE),
    avg_amenities = mean(amenities_count),
    avg_availability = mean(availability_365)
  )

# View the summary table of cluster profiles
print(mktseg_segment_summary)

```

