---
title: "cluster"
output: html_document
date: "2025-05-03"
---

```{r setup, include=FALSE}

LA_listings <- read_csv("/Users/leo/Desktop/BUS659/project/LA_listings.csv", show_col_types = FALSE)

```

The following lines are the first round of market segmenting through clustering models. Lines 15-144 are the first, less complex model, which is labelled with "general" signifiers. These are taken as simple features that are standard across Airbnbs.

```{r}

library(tidyverse)
library(cluster)
library(factoextra)
library(stringr)

# Select relevant columns and clean them for modeling, including both features of the property and indications of demand. 
mktseg_cluster_data_general <- LA_listings %>%
  select(price, bedrooms, bathrooms, beds, review_scores_rating, availability_365, amenities) %>%
  mutate(
    
    # Convert price to numeric by removing formatting
    price = as.numeric(gsub("[$,]", "", price)),
    
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    review_scores_rating = as.numeric(review_scores_rating),
    availability_365 = as.numeric(availability_365),
    amenities_count = str_count(amenities, ",") + 1
  ) %>%
  
  #Now that amenities are counted, we drop the actual amenities inputs 
  select(-amenities) %>%
  drop_na()

#Scale variables so that they contribute equally, even when initally measured by different units (price v. rating)
mktseg_cluster_data_scaled_general <- scale(mktseg_cluster_data_general)

```

```{r}

# Elbow Method to Choose Optimal K

wss <- function(k) {
  kmeans(mktseg_cluster_data_scaled_general, centers = k, nstart = 25)$tot.withinss
}
wss_values <- map_dbl(1:10, wss)

plot(1:10, wss_values, type = "b", pch = 19,
     xlab = "Number of Clusters K",
     ylab = "WSS",
     main = "Elbow Method for Optimal K")

```

```{r}

# Apply K-means clustering. Based on the elbow plot, select a k = x (adjust x to different # if appropriate)

# Add the resulting cluster assignments as a new variable,so now each row belongs to one of the k (4) market segments (labeled as 1 through 4)
# Instead of assigning directly to LA_clean (which causes a row mismatch), we assign segment labels only to the dataset used for clustering

kmeans_general_result <- kmeans(mktseg_cluster_data_scaled_general, centers = 4, nstart = 25)

fviz_cluster(kmeans_general_result, data = mktseg_cluster_data_scaled_general,
             main = "K=4 Means Clustering on Airbnb Data")

```

```{r}

# Track row IDs for consistent matching with clustering results
mktseg_clustered_general <- mktseg_cluster_data_general %>%
  mutate(segment = as.factor(kmeans_general_result$cluster))

# Summarize the characteristics of each segment, helps interpret how different segments differ in key features (e.g., price, size, amenities)
mktseg_segment_summary_general <- mktseg_clustered_general %>%
  group_by(segment) %>%
  summarise(
    avg_price = mean(price),
    avg_bedrooms = mean(bedrooms),
    avg_bathrooms = mean(bathrooms),
    avg_beds = mean(beds),
    avg_rating = mean(review_scores_rating, na.rm = TRUE),
    avg_amenities = mean(amenities_count),
    avg_availability = mean(availability_365)
  )

# View the summary table of cluster profiles
print(mktseg_segment_summary_general)

# Silhouette analysis
silhouette_score_general <- silhouette(kmeans_general_result$cluster, dist(mktseg_cluster_data_scaled_general))
plot(silhouette_score_general)

```

Understanding silhouette scores:
~0.5+ = strong structure
~0.25–0.5 = fair structure
~0.0–0.25 = weak structure, potentially overlapping clusters
< 0 = potential misclassification

Our results:

We're adding a PCA because I am doing so much guess work and I hear (read) this is helpful

```{r}

# PCA

library(ggfortify)

# Perform PCA on the scaled general clustering data
mktseg_pca_general <- prcomp(mktseg_cluster_data_scaled_general, scale = FALSE)

# Scree plot on the scaled general clustering data
screeplot(mktseg_pca_general, main = "Scree Plot of PCA")

# Base R biplot on the scaled general clustering data
biplot(mktseg_pca_general, main = "Base R Biplot")

# ggplot on the scaled general clustering data
autoplot(mktseg_pca_general, 
         data = mktseg_clustered_general, 
         colour = 'segment',
         loadings = TRUE, 
         loadings.label = TRUE, 
         loadings.label.size = 4) + 
  theme_bw() + 
  ggtitle("PCA Biplot with K-Means Segment Labels")


```

-------------------------

Lines 145-300 attempt to refine each cluster based on amenities

```{r}

library(tidyverse)
library(cluster)
library(factoextra)
library(stringr)

# Engineer binary amenities
mktseg_cluster_data_amenities <- LA_listings %>%
  select(price, bedrooms, bathrooms, beds, review_scores_rating, availability_365, amenities) %>%
  mutate(
    
    #Includes initial variables as mktseg_cluster_data_general
    price = as.numeric(gsub("[$,]", "", price)),
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    review_scores_rating = as.numeric(review_scores_rating),
    availability_365 = as.numeric(availability_365),
    amenities_count = str_count(amenities, ",") + 1,
    
    #Asked my pals in New England what they would look for on properties in Los Angeles county... I like that they asked for external cameras lol
    has_wifi = grepl("Wifi", amenities, ignore.case = TRUE),
    has_kitchen = grepl("Kitchen", amenities, ignore.case = TRUE),
    has_air_conditioning = grepl("Air conditioning", amenities, ignore.case = TRUE),
    has_external_camera = grepl("Exterior security cameras on property", amenities, ignore.case = TRUE),
    has_washer = grepl("Washer", amenities, ignore.case = TRUE),
    has_tv = grepl("TV", amenities, ignore.case = TRUE),
    has_coffee_maker = grepl("Coffee maker", amenities, ignore.case = TRUE),
    has_free_parking = grepl("Free parking", amenities, ignore.case = TRUE),
    has_private_pool = grepl("Private pool", amenities, ignore.case = TRUE)
  ) %>%
  
  select(-amenities) %>%
  
  drop_na()

# Scale features for clustering (excluding amenities_count for now)
mktseg_cluster_data_scaled_amenities <- scale(
  mktseg_cluster_data_amenities %>% select(-amenities_count)
)

```

```{r}

# Elbow method to select optimal K
wss_amenities <- function(k) {
  kmeans(mktseg_cluster_data_scaled_amenities, centers = k, nstart = 25)$tot.withinss
}
wss_values_amenities <- map_dbl(1:10, wss_amenities)

plot(1:10, wss_values_amenities, type = "b", pch = 19,
     xlab = "Number of Clusters K",
     ylab = "Total WSS",
     main = "Elbow Method for Amenities-Based Clustering")

```

```{r}

# Apply K-means clustering (adjust k as needed)
mktseg_kmeans_result_amenities <- kmeans(mktseg_cluster_data_scaled_amenities, centers = 4, nstart = 25)

fviz_cluster(mktseg_kmeans_result_amenities,
             data = mktseg_cluster_data_scaled_amenities,
             main = "K=4 Clustering Based on Binary Amenities")

```

```{r}

# Track row IDs to join segment labels back
mktseg_clustered_amenities <- mktseg_cluster_data_amenities %>%
  mutate(row_id = as.integer(rownames(.)),
         segment = as.factor(mktseg_kmeans_result_amenities$cluster))

mktseg_LA_listings_amenities <- LA_listings %>%
  mutate(row_id = row_number())

# Join clusters back to full listing data (new dataset)

mktseg_LA_segmented_amenities <- mktseg_LA_listings_amenities %>%
  left_join(select(mktseg_clustered_amenities, row_id, segment, amenities_count), by = "row_id") %>%
  select(-row_id)

# Summary of cluster profiles
mktseg_segment_summary_amenities <- mktseg_LA_segmented_amenities %>%
  filter(!is.na(segment)) %>%
  group_by(segment) %>%
  summarise(
    avg_price = mean(as.numeric(gsub("[$,]", "", price))),
    avg_bedrooms = mean(as.numeric(bedrooms)),
    avg_bathrooms = mean(as.numeric(bathrooms)),
    avg_beds = mean(as.numeric(beds)),
    avg_rating = mean(as.numeric(review_scores_rating), na.rm = TRUE),
    avg_amenities = mean(amenities_count),
    avg_availability = mean(as.numeric(availability_365))
  )

# View the summary table of cluster profiles
print(mktseg_segment_summary_amenities)

# Silhouette score

silhouette_scores_amenities <- silhouette(mktseg_kmeans_result_amenities$cluster, dist(mktseg_cluster_data_scaled_amenities))
plot(silhouette_scores_amenities)

```

```{r}

# PCA on the scaled amenities clustering data
mktseg_pca_amenities <- prcomp(mktseg_cluster_data_scaled_amenities, scale = FALSE)

# Scree plot on the scaled amenities clustering data
screeplot(mktseg_pca_amenities, main = "Scree Plot for Amenities PCA")

# Biplot on the scaled amenities clustering data
biplot(mktseg_pca_amenities, main = "Base R PCA Biplot")

# ggplot on the scaled amenities clustering data
autoplot(mktseg_pca_amenities, 
         data = mktseg_clustered_amenities,
         colour = 'segment',
         loadings = TRUE,
         loadings.label = TRUE,
         loadings.label.size = 4) +
  theme_bw() +
  ggtitle("PCA Biplot with Cluster Labels (Amenities-Based)")

```

Understanding silhouette scores:
~0.5+ = strong structure
~0.25–0.5 = fair structure
~0.0–0.25 = weak structure, potentially overlapping clusters
< 0 = potential misclassification

Our results:



Lines 290-X is an attempt to refine each cluster based on location

```{r}

# Add row_id early so we can use it later for safe joins
mktseg_cluster_data_location <- LA_listings %>%
  mutate(
    row_id = row_number(),
    is_downtown = neighbourhood_cleansed %in% c("Downtown", "Koreatown", "Arts District", "South Park"),
    is_westside = neighbourhood_cleansed %in% c("Santa Monica", "Venice", "Marina del Rey", "West LA"),
    is_eastside = neighbourhood_cleansed %in% c("Echo Park", "Silver Lake", "Los Feliz"),
    is_valley = neighbourhood_cleansed %in% c("North Hollywood", "Studio City", "Sherman Oaks", "Van Nuys")
  ) %>%
  select(row_id, price, bedrooms, bathrooms, beds, review_scores_rating, availability_365,
         is_downtown, is_westside, is_eastside, is_valley) %>%
  mutate(
    price = as.numeric(gsub("[$,]", "", price)),
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    review_scores_rating = as.numeric(review_scores_rating),
    availability_365 = as.numeric(availability_365)
  ) %>%
  drop_na()

```

```{r}

#Scale data
mktseg_cluster_data_scaled_location <- scale(mktseg_cluster_data_location %>% select(-row_id))

```

```{r}

wss_location <- function(k) {
  kmeans(mktseg_cluster_data_scaled_location, centers = k, nstart = 25)$tot.withinss
}
wss_values_location <- map_dbl(1:10, wss_location)

plot(1:10, wss_values_location, type = "b", pch = 19,
     xlab = "Number of Clusters K",
     ylab = "Total Within-Cluster Sum of Squares",
     main = "Elbow Method for Optimal K (Location Proxies)")

```

```{r}

# K-means Clustering (K = 4)
set.seed(123)
mktseg_kmeans_result_location <- kmeans(mktseg_cluster_data_scaled_location, centers = 4, nstart = 25)

fviz_cluster(mktseg_kmeans_result_location,
             data = mktseg_cluster_data_scaled_location,
             main = "K=4 Clustering Using Location Proxies")

```

```{r}

mktseg_clustered_location <- mktseg_cluster_data_location %>%
  mutate(segment = as.factor(mktseg_kmeans_result_location$cluster))

mktseg_LA_segmented_location <- mktseg_cluster_data_location %>%
  left_join(select(mktseg_clustered_location, row_id, segment), by = "row_id") %>%
  select(-row_id)

```

```{r}

mktseg_segment_summary_location <- mktseg_LA_segmented_location %>%
  filter(!is.na(segment)) %>%
  group_by(segment) %>%
  summarise(
    avg_price = mean(price),
    avg_bedrooms = mean(bedrooms),
    avg_bathrooms = mean(bathrooms),
    avg_beds = mean(beds),
    avg_rating = mean(review_scores_rating, na.rm = TRUE),
    avg_availability = mean(availability_365),
    pct_downtown = mean(is_downtown),
    pct_westside = mean(is_westside),
    pct_eastside = mean(is_eastside),
    pct_valley = mean(is_valley)
  )

print(mktseg_segment_summary_location)

silhouette_score_location <- silhouette(mktseg_kmeans_result_location$cluster, dist(mktseg_cluster_data_scaled_location))
plot(silhouette_score_location)

```

```{r}

mktseg_pca_location <- prcomp(mktseg_cluster_data_scaled_location, scale = FALSE)

screeplot(mktseg_pca_location, main = "Scree Plot (Location PCA)")
biplot(mktseg_pca_location, main = "Base R Biplot")

autoplot(mktseg_pca_location,
         data = mktseg_clustered_location,
         colour = 'segment',
         loadings = TRUE,
         loadings.label = TRUE,
         loadings.label.size = 4) +
  theme_bw() +
  ggtitle("PCA Biplot with Cluster Labels (Location-Based)")


```
