---
title: "Airbnb Data Science Project"
author: "Luke O'Connell and Yvonne Quach"
date: "May 14, 2025"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    toc_depth: 2
    theme: readable
    highlight: tango

---

# Introduction

This project applies machine learning techniques to Airbnb listings in Los Angeles County to address three core business objectives, in the hopes of providing hosts data-driven information on how best to market, price, and respond to guest demand:

1. **Price Optimization** – Determine optimal nightly prices for listings based on factors such as room type, location, amenities, and review scores.
2. **Revenue Forecasting** – Predict revenue using booking history of similar rentals, allowing hosts to anticipate seasonal demand.
3. **Market Segmentation** – Cluster listings into distinct market segments to inform business strategies, marketing efforts, and pricing tiers.

The insights aim to support both hosts and platforms in making more data-informed decisions that enhance profitability and customer satisfaction.

## Dataset Overview

## Libraries and Setup

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small

library(knitr)

opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = TRUE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```

```{r}
# Load required libraries
library(dplyr)
library(tidyr)
library(tidyverse)
library(randomForest)
library(caret)
library(stringr)
library(forcats)
library(rsample)
library(glmnet)
library(glmnetUtils)
library(tidyverse)
library(cluster)
library(factoextra)
library(stringr)
library(ggfortify)

# Load the dataset

#Yvonne, here are your load instructions: 
#LA_listings <- read_csv("LA_listings.csv", show_col_types = FALSE)

#Luke, here are your load instructions: 
LA_listings <- read_csv("/Users/leo/Desktop/BUS659/project/LA_listings.csv", show_col_types = FALSE)
```

---

# Price Optimization

Start: Price optimization... Predict optimal nightly rates based on listing’s characteristics, like room type, amenities, reviews, and location.

Since there are 3 objectives, all of which will require testing, training, and likely creation of similiar data sets, all data groups created for price optimization will begin with "opt_". 

## Data Preparation

```{r}

#Select relevant columns and clean them for modeling
opt_LA_price <- LA_listings %>% 
  select(price, room_type, neighbourhood_cleansed, number_of_reviews, review_scores_rating, minimum_nights, availability_365, latitude, longitude, amenities, bedrooms, bathrooms, beds) %>% 
  mutate(
    
    # Convert price string to numeric
    price = as.numeric(gsub("[$,]", "", price)),
    
    # Convert categorical variables to factors
    room_type = as.factor(room_type),
    neighbourhood_cleansed = as.factor(neighbourhood_cleansed),
    
    #Convert bedroom/bathroom/bed counts to numeric
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    
    # Create a new numeric feature: count of amenities
    amenities_count = str_count(amenities, ",") +1,
    
    #Price per bedroom (if bedroom is not 0), indicates price based on listing size
    price_per_bedroom = ifelse(bedrooms > 0, price / bedrooms, NA),
    
     #Price per bed, indicates price based on listing size
    price_per_bed = ifelse(beds > 0, price / beds, NA),
    
    #Indicator if listing is entire home, for price segmentation
    is_entire_home = ifelse(room_type == "Entire home/apt", 1, 0),
    
    #Log-transformed price to address skewness for sensitive distributions
    log_price = log1p(price)
  ) %>% 
  
  # Drop any rows that have missing values in the selected columns
  drop_na() %>% 
  
  # Drop the raw amenities text column 
  select(-amenities)
  
# Random Forest cannot handle factors with more than 53 levels, so we group smaller neighborhoods into an "Other" category
opt_top_neighborhoods <- opt_LA_price %>% 
  count(neighbourhood_cleansed, sort = TRUE) %>% 
  slice(1:10) %>% 
  pull(neighbourhood_cleansed) %>% 
  as.character()

opt_LA_price <- opt_LA_price %>% 
  mutate(neighbourhood_cleansed = fct_other(neighbourhood_cleansed, keep = opt_top_neighborhoods))
```

```{r}

# Boxplot of price by room type
ggplot(opt_LA_price, aes(x = room_type, y = price)) +
  geom_boxplot(fill = "lightblue") +
  labs(title = "Price Distribution by Room Type", x = "Room Type", y = "Price") +
  theme_minimal()

# Basic price map
ggplot(opt_LA_price, aes(x = longitude, y = latitude, color = price)) +
  geom_point(alpha = 0.6) +
  scale_color_viridis_c(option = "plasma") +
  labs(title = "Listings by Location and Price", x = "Longitude", y = "Latitude") +
  theme_minimal()


```


```{r}

# Split the data into training and testing sets

# Use rsample's initial_split() to create an 80/20 train/test split
set.seed(123)
opt_split <- initial_split(opt_LA_price, prop = 0.8)
    
# Extract training and testing sets
opt_LA_train <- training(opt_split)
opt_LA_test <- testing(opt_split)
 
```

```{r}

# Price distribution (before and after log transform)
ggplot(opt_LA_price, aes(x = price)) +
  geom_histogram(bins = 50, fill = "steelblue") +
  labs(title = "Distribution of Price (Raw)", x = "Price", y = "Count")

ggplot(opt_LA_price, aes(x = log_price)) +
  geom_histogram(bins = 50, fill = "darkgreen") +
  labs(title = "Distribution of Price (Log-Transformed)", x = "Log(Price + 1)", y = "Count")

```


## Model Building

### Random Forest


```{r}
# Train the Random Forest model

# Fit a random forrest regression model to predict nightly price
opt_rf_model <- randomForest(
  
  # Predice price using all other variables
  price ~ .,
  
  # Use training data only
  data = opt_LA_train,
  
  # Number of trees in the forest
  ntree = 100,
  
  # Track feature importance for interpretation
  importance = TRUE
)

# Feature importance from random forest
varImpPlot(opt_rf_model, type = 1, main = "Feature Importance (Mean Decrease Accuracy)")

```


```{r}

# Predict on the test data
predictions <- predict(opt_rf_model, newdata = opt_LA_test)

# Calculate performance metrics
mae <- mean(abs(predictions - opt_LA_test$price))
rmse <- sqrt(mean((predictions - opt_LA_test$price)^2))
r2 <- cor(predictions, opt_LA_test$price)^2

# Print the results
cat("MAE:", round(mae, 2), "\n")
cat("RMSE:", round(rmse, 2), "\n")
cat("R2:", round(r2, 3), "\n")

# Actual vs Predicted price scatterplot
ggplot(data.frame(actual = opt_LA_test$price, predicted = predictions), aes(x = actual, y = predicted)) +
  geom_point(alpha = 0.4) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Actual vs Predicted Prices", x = "Actual Price", y = "Predicted Price") +
  theme_minimal()

```

### Another modeling option?

## Model Evaluation

MAE: 4.05 
RMSE: 83.49 
R2: 0.975 

The R2 result is very high, and explains 97.5% of the variance seen in Airbnb prices. The MAE is relatively low and, like R2, suggests that the model performs well for most listings. The high RSME indicates that there are a few extreme-priced listings that produce large errors in the dataset. These listings may include luxury properties or otherwise unique offerings. 

Do not delete, reference draft 1 (?) for initial model and compare this analysis to our final analysis: The MAE is $ 69.52, so on average, our predictions are off by $70 per listing, which is a slight improvement from before. The RMSE is $273.42, so our model still has some large errors on more expensive listings, but slightly better than before. The r-squared is 0.587, which means our model now explains 59% of the variance in Airbnb prices, a slight improvement from 56%. 

## Business Implications

Further training can be done to address the skewed RMSE, potentially isolating the higher-end listings separately. Airbnb may want to act on this separate market segment and offer hosts different marketing, services, and other opportunities to better position their properties to target audiences. 

Since this model explains 97.5% of variance in pricing, this can provide information for hosts to set competitive prices based on location or attributes rather than intuition. This will be espicially valuable for hosts introducing new listings who can more imediately evaluate their property instead of trial and error. Airbnb can act on this data by recommending nighltly prices to hosts that align with market trends.

This also may encourage hosts to introduce new featrues to their rentals, if they are looking for an opportunity to charge more or become competitive. Airbnb can offer an ROI calculator for upgrades to their properties, such as expanding amenties or adding a couch to increase capacity, and the return on that investment. 

---

# Revenue Forecasting

Start: Revenue forecasting... Estimate annual revenue per listing using reported attributes of each listing for hosts to make informed decisions about setting prices and availability on their properties. Estimate expected future revenue for listings using pricing, availability, and booking history.

Since there are 3 objectives, all of which will require testing, training, and likely creation of similar data sets, all data groups created for price optimization will begin with "rev_". 

## Data Preparation

```{r}

# Select relevant columns and clean them for modeling

rev_LA_avail <- LA_listings %>%
  select(price, room_type, neighbourhood_cleansed, number_of_reviews, review_scores_rating,
         minimum_nights, availability_365, latitude, longitude, amenities,
         bedrooms, bathrooms, beds) %>%
  mutate(
    # Convert price string to numeric
    price = as.numeric(gsub("[$,]", "", price)),
    
    # Create proxy for bookings, allowing number of reviews to stand in for bookings if no booking data is available
    #This assumes a correlation between reviews and bookings, which likely underestimates revenue for newer listings
    proxy_bookings = number_of_reviews,
    
    # Create revenue variable and log-transformed revenue
    revenue = price * proxy_bookings,
    log_revenue = log1p(revenue),
    
    # Isolate availability by singular days
    availability_rate = availability_365 / 365,
    
    # Convert categorical variables to factors
    room_type = as.factor(room_type),
    neighbourhood_cleansed = as.factor(neighbourhood_cleansed),
    
    # Convert bedroom/bathroom/bed counts to numeric
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    
    # Create a new numeric feature: count of amenities
    amenities_count = str_count(amenities, ",") + 1
  ) %>%
  select(-amenities) %>%  # Drop raw amenities text
  drop_na()

par(mfrow = c(1, 2))
hist(rev_LA_avail$revenue, main = "Raw Revenue", xlab = "Revenue", breaks = 50, col = "lightblue")
hist(rev_LA_avail$log_revenue, main = "Log Revenue", xlab = "Log(1 + Revenue)", breaks = 50, col = "lightgreen")

```

```{r}

# Create an 80/20 train/test split

rev_split <- initial_split(rev_LA_avail, prop = 0.8)

# Extract training and testing sets
rev_train <- training(rev_split)
rev_test  <- testing(rev_split)

boxplot(log_revenue ~ room_type, data = rev_LA_avail,
        col = "lightgray", main = "Log Revenue by Room Type",
        xlab = "Room Type", ylab = "Log(Revenue + 1)")

```

## Data Engineering

```{r}

# We did not learn model.matrix in class, however, it was covered in BDS's section on regression and model selection in chapter 2

# Create revenue training and testing matrices
rev_train_matrix <- model.matrix(~ . -1, data = rev_train %>% select(-log_revenue))
rev_train_target <- rev_train$log_revenue

rev_test_matrix <- model.matrix(~ . -1, data = rev_test %>% select(-log_revenue))
rev_test_target <- rev_test$log_revenue

```

## Forecasting Models

### Elastic Net

```{r}

# Fit by using cross-validation across multiple alpha values
set.seed(123)

enet_mod <- cva.glmnet(
  x = rev_train_matrix,
  y = rev_train_target,
  alpha = seq(0, 1, length = 11),
  family = "gaussian"
)

# I pulled this from regularization slide page 55. Helper function to find the alpha with lowest cross-validation error
rev_get_alpha <- function(fit) {
  alpha <- fit$alpha
  error <- sapply(fit$modlist, function(mod) {min(mod$cvm)})
  alpha[which.min(error)]
}

# Helper function to extract alpha, lambda.min, lambda.1se, and error
rev_get_model_params <- function(fit) {
  alpha <- fit$alpha
  lambdaMin <- sapply(fit$modlist, `[[`, "lambda.min")
  lambdaSE <- sapply(fit$modlist, `[[`, "lambda.1se")
  error <- sapply(fit$modlist, function(mod) {min(mod$cvm)})
  best <- which.min(error)
  data.frame(alpha = alpha[best], lambdaMin = lambdaMin[best], lambdaSE = lambdaSE[best], error = error[best])
}

# Identify best alpha and lambda values
rev_best_alpha <- rev_get_alpha(enet_mod)
print(paste("Best alpha:", rev_best_alpha))

rev_model_params <- rev_get_model_params(enet_mod)
print(rev_model_params)

# Extract best model using optimal alpha
rev_best_model <- enet_mod$modlist[[which(enet_mod$alpha == rev_best_alpha)]]

# Predict log revenue using the best model and optimal lambda
rev_enet_predictions <- predict(
  rev_best_model,
  s = rev_model_params$lambdaMin,
  newx = rev_test_matrix
)

# Evaluate model performance on the test set
rev_enet_mae <- mean(abs(rev_enet_predictions - rev_test_target))
rev_enet_rmse <- sqrt(mean((rev_enet_predictions - rev_test_target)^2))
rev_enet_r2 <- cor(rev_enet_predictions, rev_test_target)^2

# Output performance results
cat("Elastic Net (best alpha =", rev_best_alpha, ") - MAE:", round(rev_enet_mae, 2), "\n")
cat("Elastic Net (best alpha =", rev_best_alpha, ") - RMSE:", round(rev_enet_rmse, 2), "\n")
cat("Elastic Net (best alpha =", rev_best_alpha, ") - R²:", round(rev_enet_r2, 3), "\n")

plot(rev_test_target, rev_enet_predictions, 
     xlab = "Actual Log Revenue", ylab = "Predicted Log Revenue",
     main = "Elastic Net Predictions", col = "darkgreen", pch = 20)
abline(0, 1, col = "red", lwd = 2)

```

## Model Evaluation

MAE: 0.95 
RMSE: 1.16 
R2: 0.616 

Since we are unable to see actual revenue per listing, and it can be assumed that revenue is highly skewed based on seasonality and dynamic pricing, most of these numbers are fine for now. We should fine tune them. We may try adding superhost status, property_type?

We may also want to consider removing log values... If we want to see actual dollar value of revenue. Luke will mess with that later. 

End: Revenue forecasting

---

# Market Segmentation

## General Segmentation

The following lines are the first round of market segmenting through clustering models. Lines 15-144 are the first, less complex model, which is labelled with "general" signifiers. These are taken as simple features that are standard across Airbnbs.

### Data Preperation

```{r}

# Select relevant columns and clean them for modeling, including both features of the property and indications of demand. 
mktseg_cluster_data_general <- LA_listings %>%
  select(price, bedrooms, bathrooms, beds, review_scores_rating, availability_365, amenities) %>%
  mutate(
    
    # Convert price to numeric by removing formatting
    price = as.numeric(gsub("[$,]", "", price)),
    
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    review_scores_rating = as.numeric(review_scores_rating),
    availability_365 = as.numeric(availability_365),
    amenities_count = str_count(amenities, ",") + 1
  ) %>%
  
  #Now that amenities are counted, we drop the actual amenities inputs 
  select(-amenities) %>%
  drop_na()

#Scale variables so that they contribute equally, even when initally measured by different units (price v. rating)
mktseg_cluster_data_scaled_general <- scale(mktseg_cluster_data_general)

```

### Modeling

#### K-means

```{r}

# Elbow Method to Choose Optimal K

wss <- function(k) {
  kmeans(mktseg_cluster_data_scaled_general, centers = k, nstart = 25)$tot.withinss
}
wss_values <- map_dbl(1:10, wss)

plot(1:10, wss_values, type = "b", pch = 19,
     xlab = "Number of Clusters K",
     ylab = "WSS",
     main = "Elbow Method for Optimal K")

```

```{r}

# Apply K-means clustering. Based on the elbow plot, select a k = x (adjust x to different # if appropriate)

# Add the resulting cluster assignments as a new variable,so now each row belongs to one of the k (4) market segments (labeled as 1 through 4)
# Instead of assigning directly to LA_clean (which causes a row mismatch), we assign segment labels only to the dataset used for clustering

kmeans_general_result <- kmeans(mktseg_cluster_data_scaled_general, centers = 4, nstart = 25)

fviz_cluster(kmeans_general_result, data = mktseg_cluster_data_scaled_general,
             main = "K=4 Means Clustering on Airbnb Data")

```

```{r}

# Track row IDs for consistent matching with clustering results
mktseg_clustered_general <- mktseg_cluster_data_general %>%
  mutate(segment = as.factor(kmeans_general_result$cluster))

# Summarize the characteristics of each segment, helps interpret how different segments differ in key features (e.g., price, size, amenities)
mktseg_segment_summary_general <- mktseg_clustered_general %>%
  group_by(segment) %>%
  summarise(
    avg_price = mean(price),
    avg_bedrooms = mean(bedrooms),
    avg_bathrooms = mean(bathrooms),
    avg_beds = mean(beds),
    avg_rating = mean(review_scores_rating, na.rm = TRUE),
    avg_amenities = mean(amenities_count),
    avg_availability = mean(availability_365)
  )

# View the summary table of cluster profiles
print(mktseg_segment_summary_general)

# Silhouette analysis
silhouette_score_general <- silhouette(kmeans_general_result$cluster, dist(mktseg_cluster_data_scaled_general))
plot(silhouette_score_general)

```

Understanding silhouette scores:
~0.5+ = strong structure
~0.25–0.5 = fair structure
~0.0–0.25 = weak structure, potentially overlapping clusters
< 0 = potential misclassification

Our results:

#### PCA 

We're adding a PCA because I am doing so much guess work and I hear (read) this is helpful

```{r}

# PCA

# Perform PCA on the scaled general clustering data
mktseg_pca_general <- prcomp(mktseg_cluster_data_scaled_general, scale = FALSE)

# Scree plot on the scaled general clustering data
screeplot(mktseg_pca_general, main = "Scree Plot of PCA")

# Base R biplot on the scaled general clustering data
biplot(mktseg_pca_general, main = "Base R Biplot")

# ggplot on the scaled general clustering data
autoplot(mktseg_pca_general, 
         data = mktseg_clustered_general, 
         colour = 'segment',
         loadings = TRUE, 
         loadings.label = TRUE, 
         loadings.label.size = 4) + 
  theme_bw() + 
  ggtitle("PCA Biplot with K-Means Segment Labels")


```

-------------------------

## Amenities Segmentation

### Data Preparation

```{r}

# Engineer binary amenities
mktseg_cluster_data_amenities <- LA_listings %>%
  select(price, bedrooms, bathrooms, beds, review_scores_rating, availability_365, amenities) %>%
  mutate(
    
    #Includes initial variables as mktseg_cluster_data_general
    price = as.numeric(gsub("[$,]", "", price)),
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    review_scores_rating = as.numeric(review_scores_rating),
    availability_365 = as.numeric(availability_365),
    amenities_count = str_count(amenities, ",") + 1,
    
    #Asked my pals in New England what they would look for on properties in Los Angeles county... I like that they asked for external cameras lol
    has_wifi = grepl("Wifi", amenities, ignore.case = TRUE),
    has_kitchen = grepl("Kitchen", amenities, ignore.case = TRUE),
    has_air_conditioning = grepl("Air conditioning", amenities, ignore.case = TRUE),
    has_external_camera = grepl("Exterior security cameras on property", amenities, ignore.case = TRUE),
    has_washer = grepl("Washer", amenities, ignore.case = TRUE),
    has_tv = grepl("TV", amenities, ignore.case = TRUE),
    has_coffee_maker = grepl("Coffee maker", amenities, ignore.case = TRUE),
    has_free_parking = grepl("Free parking", amenities, ignore.case = TRUE),
    has_private_pool = grepl("Private pool", amenities, ignore.case = TRUE)
  ) %>%
  
  select(-amenities) %>%
  
  drop_na()

# Scale features for clustering (excluding amenities_count for now)
mktseg_cluster_data_scaled_amenities <- scale(
  mktseg_cluster_data_amenities %>% select(-amenities_count)
)

```

### Modeling

#### K-means

```{r}

# Elbow method to select optimal K
wss_amenities <- function(k) {
  kmeans(mktseg_cluster_data_scaled_amenities, centers = k, nstart = 25)$tot.withinss
}
wss_values_amenities <- map_dbl(1:10, wss_amenities)

plot(1:10, wss_values_amenities, type = "b", pch = 19,
     xlab = "Number of Clusters K",
     ylab = "Total WSS",
     main = "Elbow Method for Amenities-Based Clustering")

```

```{r}

# Apply K-means clustering (adjust k as needed)
mktseg_kmeans_result_amenities <- kmeans(mktseg_cluster_data_scaled_amenities, centers = 4, nstart = 25)

fviz_cluster(mktseg_kmeans_result_amenities,
             data = mktseg_cluster_data_scaled_amenities,
             main = "K=4 Clustering Based on Binary Amenities")

ggplot(mktseg_kmeans_result_amenities, aes(x = segment, y = avg_amenities)) +
  geom_col(fill = "darkorange") +
  labs(title = "Avg Amenities by Segment", x = "Segment", y = "Average Amenity Count")


```

```{r}

# Track row IDs to join segment labels back
mktseg_clustered_amenities <- mktseg_cluster_data_amenities %>%
  mutate(row_id = as.integer(rownames(.)),
         segment = as.factor(mktseg_kmeans_result_amenities$cluster))

mktseg_LA_listings_amenities <- LA_listings %>%
  mutate(row_id = row_number())

# Join clusters back to full listing data (new dataset)

mktseg_LA_segmented_amenities <- mktseg_LA_listings_amenities %>%
  left_join(select(mktseg_clustered_amenities, row_id, segment, amenities_count), by = "row_id") %>%
  select(-row_id)

# Summary of cluster profiles
mktseg_segment_summary_amenities <- mktseg_LA_segmented_amenities %>%
  filter(!is.na(segment)) %>%
  group_by(segment) %>%
  summarise(
    avg_price = mean(as.numeric(gsub("[$,]", "", price))),
    avg_bedrooms = mean(as.numeric(bedrooms)),
    avg_bathrooms = mean(as.numeric(bathrooms)),
    avg_beds = mean(as.numeric(beds)),
    avg_rating = mean(as.numeric(review_scores_rating), na.rm = TRUE),
    avg_amenities = mean(amenities_count),
    avg_availability = mean(as.numeric(availability_365))
  )

# View the summary table of cluster profiles
print(mktseg_segment_summary_amenities)

# Silhouette score

silhouette_scores_amenities <- silhouette(mktseg_kmeans_result_amenities$cluster, dist(mktseg_cluster_data_scaled_amenities))
plot(silhouette_scores_amenities)

```

```{r}

library(ggplot2)
ggplot(mktseg_segment_summary_general, aes(x = segment, y = avg_price)) +
  geom_col(fill = "steelblue") +
  labs(title = "Avg Price by Segment", x = "Segment", y = "Average Price")


```

#### PCA 

```{r}

# PCA on the scaled amenities clustering data
mktseg_pca_amenities <- prcomp(mktseg_cluster_data_scaled_amenities, scale = FALSE)

# Scree plot on the scaled amenities clustering data
screeplot(mktseg_pca_amenities, main = "Scree Plot for Amenities PCA")

# Biplot on the scaled amenities clustering data
biplot(mktseg_pca_amenities, main = "Base R PCA Biplot")

# ggplot on the scaled amenities clustering data
autoplot(mktseg_pca_amenities, 
         data = mktseg_clustered_amenities,
         colour = 'segment',
         loadings = TRUE,
         loadings.label = TRUE,
         loadings.label.size = 4) +
  theme_bw() +
  ggtitle("PCA Biplot with Cluster Labels (Amenities-Based)")

```

Understanding silhouette scores:
~0.5+ = strong structure
~0.25–0.5 = fair structure
~0.0–0.25 = weak structure, potentially overlapping clusters
< 0 = potential misclassification

Our results:

## Location Segmentation

### Data Preparation

```{r}

# Add row_id early so we can use it later for safe joins
mktseg_cluster_data_location <- LA_listings %>%
  mutate(
    row_id = row_number(),
    is_downtown = neighbourhood_cleansed %in% c("Downtown", "Koreatown", "Arts District", "South Park"),
    is_westside = neighbourhood_cleansed %in% c("Santa Monica", "Venice", "Marina del Rey", "West LA"),
    is_eastside = neighbourhood_cleansed %in% c("Echo Park", "Silver Lake", "Los Feliz"),
    is_valley = neighbourhood_cleansed %in% c("North Hollywood", "Studio City", "Sherman Oaks", "Van Nuys")
  ) %>%
  select(row_id, price, bedrooms, bathrooms, beds, review_scores_rating, availability_365,
         is_downtown, is_westside, is_eastside, is_valley) %>%
  mutate(
    price = as.numeric(gsub("[$,]", "", price)),
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    review_scores_rating = as.numeric(review_scores_rating),
    availability_365 = as.numeric(availability_365)
  ) %>%
  drop_na()

```

```{r}

#Scale data
mktseg_cluster_data_scaled_location <- scale(mktseg_cluster_data_location %>% select(-row_id))

```

### Modeling

#### K-means

```{r}

wss_location <- function(k) {
  kmeans(mktseg_cluster_data_scaled_location, centers = k, nstart = 25)$tot.withinss
}
wss_values_location <- map_dbl(1:10, wss_location)

plot(1:10, wss_values_location, type = "b", pch = 19,
     xlab = "Number of Clusters K",
     ylab = "Total Within-Cluster Sum of Squares",
     main = "Elbow Method for Optimal K (Location Proxies)")

```

```{r}

# K-means Clustering (K = 4)
set.seed(123)
mktseg_kmeans_result_location <- kmeans(mktseg_cluster_data_scaled_location, centers = 4, nstart = 25)

fviz_cluster(mktseg_kmeans_result_location,
             data = mktseg_cluster_data_scaled_location,
             main = "K=4 Clustering Using Location Proxies")

```

```{r}

mktseg_clustered_location <- mktseg_cluster_data_location %>%
  mutate(segment = as.factor(mktseg_kmeans_result_location$cluster))

mktseg_LA_segmented_location <- mktseg_cluster_data_location %>%
  left_join(select(mktseg_clustered_location, row_id, segment), by = "row_id") %>%
  select(-row_id)

```

```{r}

mktseg_segment_summary_location <- mktseg_LA_segmented_location %>%
  filter(!is.na(segment)) %>%
  group_by(segment) %>%
  summarise(
    avg_price = mean(price),
    avg_bedrooms = mean(bedrooms),
    avg_bathrooms = mean(bathrooms),
    avg_beds = mean(beds),
    avg_rating = mean(review_scores_rating, na.rm = TRUE),
    avg_availability = mean(availability_365),
    pct_downtown = mean(is_downtown),
    pct_westside = mean(is_westside),
    pct_eastside = mean(is_eastside),
    pct_valley = mean(is_valley)
  )

print(mktseg_segment_summary_location)

silhouette_score_location <- silhouette(mktseg_kmeans_result_location$cluster, dist(mktseg_cluster_data_scaled_location))
plot(silhouette_score_location)

```

#### PCA 


```{r}

mktseg_pca_location <- prcomp(mktseg_cluster_data_scaled_location, scale = FALSE)

screeplot(mktseg_pca_location, main = "Scree Plot (Location PCA)")
biplot(mktseg_pca_location, main = "Base R Biplot")

autoplot(mktseg_pca_location,
         data = mktseg_clustered_location,
         colour = 'segment',
         loadings = TRUE,
         loadings.label = TRUE,
         loadings.label.size = 4) +
  theme_bw() +
  ggtitle("PCA Biplot with Cluster Labels (Location-Based)")

location_pct_long <- mktseg_segment_summary_location %>%
  pivot_longer(cols = starts_with("pct_"), names_to = "region", values_to = "percentage")

ggplot(location_pct_long, aes(x = segment, y = percentage, fill = region)) +
  geom_col(position = "dodge") +
  labs(title = "Proportion of Region by Segment", x = "Segment", y = "Proportion") +
  scale_y_continuous(labels = scales::percent_format())

```

---

# Conclusion

## Summary of Findings

## Limitations

## Recommendations for Future Work

