---
title: "Airbnb"
author: "Luke O'Connell & Yvonne Quach"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small

library(knitr)

opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = TRUE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```



```{r}
# Load required libraries
library(dplyr)
library(tidyr)
library(tidyverse)
library(randomForest)
library(caret)
library(stringr)
library(forcats)

# Load the dataset

#Yvonne, here are your load instructions: 
#LA_listings <- read_csv("LA_listings.csv", show_col_types = FALSE)

#Luke, here are your load instructions: 
LA_listings <- read_csv("/Users/leo/Desktop/BUS659/project/LA_listings.csv", show_col_types = FALSE)
```

Start: Price optimization

Since there are 3 objectives, all of which will require testing, training, and likely creation of similiar data sets, all data groups created for price optimization will begin with "opt_". 

```{r}

#Select relevant columns and clean them for modeling
opt_LA_price <- LA_listings %>% 
  select(price, room_type, neighbourhood_cleansed, number_of_reviews, review_scores_rating, minimum_nights, availability_365, latitude, longitude, amenities, bedrooms, bathrooms, beds) %>% 
  mutate(
    
    # Convert price string to numeric
    price = as.numeric(gsub("[$,]", "", price)),
    
    # Convert categorical variables to factors
    room_type = as.factor(room_type),
    neighbourhood_cleansed = as.factor(neighbourhood_cleansed),
    
    #Convert bedroom/bathroom/bed counts to numeric
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    
    # Create a new numeric feature: count of amenities
    amenities_count = str_count(amenities, ",") +1,
    
    #Price per bedroom (if bedroom is not 0), indicates price based on listing size
    price_per_bedroom = ifelse(bedrooms > 0, price / bedrooms, NA),
    
     #Price per bed, indicates price based on listing size
    price_per_bed = ifelse(beds > 0, price / beds, NA),
    
    #Indicator if listing is entire home, for price segmentation
    is_entire_home = ifelse(room_type == "Entire home/apt", 1, 0),
    
    #Log-transformed price to address skewness for sensitive distributions
    log_price = log1p(price)
  ) %>% 
  
  # Drop any rows that have missing values in the selected columns
  drop_na() %>% 
  
  # Drop the raw amenities text column 
  select(-amenities)
  
# Random Forest cannot handle factors with more than 53 levels, so we group smaller neighborhoods into an "Other" category
opt_top_neighborhoods <- opt_LA_price %>% 
  count(neighbourhood_cleansed, sort = TRUE) %>% 
  slice(1:10) %>% 
  pull(neighbourhood_cleansed) %>% 
  as.character()

opt_LA_price <- opt_LA_price %>% 
  mutate(neighbourhood_cleansed = fct_other(neighbourhood_cleansed, keep = opt_top_neighborhoods))
```


```{r}

# Split the data into training and testing sets

# Load rsample
library(rsample)

# Use rsample's initial_split() to create an 80/20 train/test split
set.seed(123)
opt_split <- initial_split(opt_LA_price, prop = 0.8)
    
# Extract training and testing sets
opt_LA_train <- training(opt_split)
opt_LA_test <- testing(opt_split)
 
```

```{r}
# Train the Random Forest model

# Fit a random forrest regression model to predict nightly price
opt_rf_model <- randomForest(
  
  # Predice price using all other variables
  price ~ .,
  
  # Use training data only
  data = opt_LA_train,
  
  # Number of trees in the forest
  ntree = 100,
  
  # Track feature importance for interpretation
  importance = TRUE
)

```


```{r}

# Predict on the test data
predictions <- predict(opt_rf_model, newdata = opt_LA_test)

# Calculate performance metrics
mae <- mean(abs(predictions - opt_LA_test$price))
rmse <- sqrt(mean((predictions - opt_LA_test$price)^2))
r2 <- cor(predictions, opt_LA_test$price)^2

# Print the results
cat("MAE:", round(mae, 2), "\n")
cat("RMSE:", round(rmse, 2), "\n")
cat("R2:", round(r2, 3), "\n")

```
MAE: 4.05 (much lower error)
RMSE: 83.49 (lower is better... not perfect, but better)
R2: 0.975 (explains 0.415 more of the variance)

```{r}

#The new MAE is $69.52, so on average, our predictions are off by $70 per listing, which is a slight improvement from before. The RMSE is $273.42, so our model still has some large errors on more expensive listings, but slightly better than before. The r-squared is 0.587, which means our model now explains 59% of the variance in Airbnb prices, a slight improvement from 56%. 
```

End: Price optimization

Start: Revenue forecasting... Estimate annual revenue per listing using reported attributes of each listing for hosts to make informed decisions about setting prices and availability on their properties. 

Since there are 3 objectives, all of which will require testing, training, and likely creation of similar data sets, all data groups created for price optimization will begin with "rev_". 

```{r}

# Select relevant columns and clean them for modeling

rev_LA_avail <- LA_listings %>%
  select(price, room_type, neighbourhood_cleansed, number_of_reviews, review_scores_rating,
         minimum_nights, availability_365, latitude, longitude, amenities,
         bedrooms, bathrooms, beds) %>%
  mutate(
    # Convert price string to numeric
    price = as.numeric(gsub("[$,]", "", price)),
    
    # Create proxy, allowing number of reviews to stand in for bookings if no booking data is available
    proxy_bookings = number_of_reviews,
    
    # Create revenue variable and log-transformed revenue
    revenue = price * proxy_bookings,
    log_revenue = log1p(revenue),
    
    # Isolate availability by singular days
    availability_rate = availability_365 / 365,
    
    # Convert categorical variables to factors
    room_type = as.factor(room_type),
    neighbourhood_cleansed = as.factor(neighbourhood_cleansed),
    
    # Convert bedroom/bathroom/bed counts to numeric
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    
    # Create a new numeric feature: count of amenities
    amenities_count = str_count(amenities, ",") + 1
  ) %>%
  select(-amenities) %>%  # Drop raw amenities text
  drop_na()
```

```{r}

# Create an 80/20 train/test split

rev_split <- initial_split(rev_LA_avail, prop = 0.8)

# Extract training and testing sets
rev_train <- training(rev_split)
rev_test  <- testing(rev_split)

```

```{r}

# Load additional required libraries for Elastic Net

library(glmnet)
library(glmnetUtils)

# We did not learn model.matrix in class, however, it was covered in BDS's section on regression and model selection in chapter 2

# Create revenue training and testing matrices
rev_train_matrix <- model.matrix(~ . -1, data = rev_train %>% select(-log_revenue))
rev_train_target <- rev_train$log_revenue

rev_test_matrix <- model.matrix(~ . -1, data = rev_test %>% select(-log_revenue))
rev_test_target <- rev_test$log_revenue

```

```{r}

# Fit by using cross-validation across multiple alpha values
set.seed(123)

enet_mod <- cva.glmnet(
  x = rev_train_matrix,
  y = rev_train_target,
  alpha = seq(0, 1, length = 11),
  family = "gaussian"
)

# I pulled this from regularization slide page 55. Helper function to find the alpha with lowest cross-validation error
rev_get_alpha <- function(fit) {
  alpha <- fit$alpha
  error <- sapply(fit$modlist, function(mod) {min(mod$cvm)})
  alpha[which.min(error)]
}

# Helper function to extract alpha, lambda.min, lambda.1se, and error
rev_get_model_params <- function(fit) {
  alpha <- fit$alpha
  lambdaMin <- sapply(fit$modlist, `[[`, "lambda.min")
  lambdaSE <- sapply(fit$modlist, `[[`, "lambda.1se")
  error <- sapply(fit$modlist, function(mod) {min(mod$cvm)})
  best <- which.min(error)
  data.frame(alpha = alpha[best], lambdaMin = lambdaMin[best], lambdaSE = lambdaSE[best], error = error[best])
}

# Identify best alpha and lambda values
rev_best_alpha <- rev_get_alpha(enet_mod)
print(paste("Best alpha:", rev_best_alpha))

rev_model_params <- rev_get_model_params(enet_mod)
print(rev_model_params)

# Extract best model using optimal alpha
rev_best_model <- enet_mod$modlist[[which(enet_mod$alpha == rev_best_alpha)]]

# Predict log revenue using the best model and optimal lambda
rev_enet_predictions <- predict(
  rev_best_model,
  s = rev_model_params$lambdaMin,
  newx = rev_test_matrix
)

# Evaluate model performance on the test set
rev_enet_mae <- mean(abs(rev_enet_predictions - rev_test_target))
rev_enet_rmse <- sqrt(mean((rev_enet_predictions - rev_test_target)^2))
rev_enet_r2 <- cor(rev_enet_predictions, rev_test_target)^2

# Output performance results
cat("Elastic Net (best alpha =", rev_best_alpha, ") - MAE:", round(rev_enet_mae, 2), "\n")
cat("Elastic Net (best alpha =", rev_best_alpha, ") - RMSE:", round(rev_enet_rmse, 2), "\n")
cat("Elastic Net (best alpha =", rev_best_alpha, ") - R²:", round(rev_enet_r2, 3), "\n")
```
