---
title: "Airbnb"
author: "Luke O'Connell & Yvonne Quach"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---

```{r setup, include=FALSE}

# Please leave this code chunk as is. It makes some slight formatting changes to alter the output to be more aesthetically pleasing. 

library(knitr)

# Change the number in set seed to your own favorite number
set.seed(1818)
options(width=70)
options(scipen=99)


# this sets text outputted in code chunks to small
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=TRUE, size = "vsmall")  
opts_chunk$set(message = FALSE,                                          
               warning = FALSE,
               # "caching" stores objects in code chunks and only rewrites if you change things
               cache = TRUE,                               
               # automatically downloads dependency files
               autodep = TRUE,
               # 
               cache.comments = FALSE,
               # 
               collapse = TRUE,
               # change fig.width and fig.height to change the code height and width by default
               fig.width = 5.5,  
               fig.height = 4.5,
               fig.align='center')


```

```{r setup-2}

# Always print this out before your assignment
sessionInfo()
getwd()

```


<!-- ### start answering your problem set here -->
<!-- You may export your homework in either html or pdf, with the former usually being easier. 
     To export or compile your Rmd file: click above on 'Knit' then 'Knit to HTML' -->
<!-- Be sure to submit both your .Rmd file and the compiled .html or .pdf file for full credit -->


```{r setup-3}

# load all your libraries in this chunk 
library('dplyr')
# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```



## Question 1

1a) 

```{r}
# Load required libraries
library(tidyverse)
library(randomForest)
library(caret)
library(stringr)
library(forcats)

# Load the dataset
LA_listings <- read_csv("LA_listings.csv", show_col_types = FALSE)

# Select and clean relevant columns
#Select relevant columns and clean them for modeling
LA_clean <- LA_listings %>% 
  select(price, room_type, neighbourhood_cleansed, number_of_reviews, review_scores_rating, minimum_nights, availability_365, latitude, longitude, amenities, bedrooms, bathrooms, beds) %>% 
  mutate(
    
    # Convert price string to numeric
    price = as.numeric(gsub("[$,]", "", price)),
    
    # Convert categorical variables to factors
    room_type = as.factor(room_type),
    neighbourhood_cleansed = as.factor(neighbourhood_cleansed),
    
    #Convert bedroom/bathroom/bed counts to numeric
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    
    # Create a new numeric feature: count of amenities
    amenities_count = str_count(amenities, ",") +1
  ) %>% 
  
  # Drop any rows that have missing values in the selected columns
  drop_na() %>% 
  
  # Drop the raw amenities text column 
  select(-amenities)
  
# Random Forest cannot handle factors with more than 53 levels, so we group smaller neighborhoods into an "Other" category
top_neighborhoods <- LA_clean %>% 
  count(neighbourhood_cleansed, sort = TRUE) %>% 
  slice(1:10) %>% 
  pull(neighbourhood_cleansed) %>% 
  as.character()

LA_clean <- LA_clean %>% 
  mutate(neighbourhood_cleansed = fct_other(neighbourhood_cleansed, keep = top_neighborhoods))
```


1b) 

```{r}

# Split the data into training and testing sets

# Load rsample
library(rsample)

# Use rsample's initial_split() to create an 80/20 train/test split
set.seed(123)
split <- initial_split(LA_clean, prop = 0.8)
    
# Extract training and testing sets
LA_train <- training(split)
LA_test <- testing(split)
 
```



1c) 

```{r}
# Train the Random Forest model

# Fit a random forrest regression model to predict nightly price
rf_model <- randomForest(
  
  # Predice price using all other variables
  price ~ .,
  
  # Use training data only
  data = LA_train,
  
  # Number of trees in the forest
  ntree = 100,
  
  # Track feature importance for interpretation
  importance = TRUE
)

```


1d) We got a Mean Absolute Error (MAE) of 78.44, which means on average, our predictions are off by about $78 per night. We got a Root Mean Squared Error (RMSE) of 276.04, which is high and could suggests some listings are priced very differently than what our model predicts. This could be due to luxury/high-end listings that can skew our RMSE and any outliers that are unusually priced. We got an R-Squared of 0.56, which means our model explains about 56% of the variance in price, which is not bad, but could use improvement. 

```{r}

# Predict on the test data
predictions <- predict(rf_model, newdata = LA_test)

# Calculate performance metrics
mae <- mean(abs(predictions - LA_test$price))
rmse <- sqrt(mean((predictions - LA_test$price)^2))
r2 <- cor(predictions, LA_test$price)^2

# Print the results
cat("MAE:", round(mae, 2), "\n")
cat("RMSE:", round(rmse, 2), "\n")
cat("R2:", round(r2, 3), "\n")

```


1e) Response to part b. 

```{r}

# We could improve our model by log_price

# Update LA-Clean code block to include log_price
LA_clean <- LA_listings %>% 
  select(price, room_type, neighbourhood_cleansed, number_of_reviews, review_scores_rating, minimum_nights, availability_365, latitude, longitude, amenities, bedrooms, bathrooms, beds) %>% 
  mutate(
    price = as.numeric(gsub("[$,]", "", price)),
    room_type = as.factor(room_type),
    neighbourhood_cleansed = as.factor(neighbourhood_cleansed),
    bedrooms = as.numeric(bedrooms),
    bathrooms = as.numeric(bathrooms),
    beds = as.numeric(beds),
    amenities_count = str_count(amenities, ",") + 1,
    log_price = log1p(price)
  ) %>% 
  drop_na() %>% 
  select(-amenities)
  
top_neighborhoods <- LA_clean %>% 
  count(neighbourhood_cleansed, sort = TRUE) %>% 
  slice(1:10) %>% 
  pull(neighbourhood_cleansed) %>% 
  as.character()

LA_clean <- LA_clean %>% 
  mutate(neighbourhood_cleansed = fct_other(neighbourhood_cleansed, keep = top_neighborhoods))
```


1f) Response to part b. 

```{r}

# Now we split the data again
set.seed(123)
split <- initial_split(LA_clean, prop = 0.8)
LA_train <- training(split)
LA_test <- testing(split)

```


1g) Response to part b. 

```{r}
# Train the model using log_price

rf_model1 <- randomForest(
  log_price ~. -price,
  data = LA_train,
  ntree = 100,
  importance = TRUE
)


```


1h) The new MAE is $69.52, so on average, our predictions are off by $70 per listing, which is a slight improvement from before. The RMSE is $273.42, so our model still has some large errors on more expensive listings, but slightly better than before. The r-squared is 0.587, which means our model now explains 59% of the variance in Airbnb prices, a slight improvement from 56%. 

```{r}

# Predict test data
log_preds <- predict(rf_model1, newdata = LA_test)
preds <- expm1(log_preds)

# Evaluate predictions against actual price
mae <- mean(abs(preds - LA_test$price))
rmse <- sqrt(mean((preds - LA_test$price)^2))
r2 <- cor(preds, LA_test$price)^2

cat("MAE:", round(mae, 2), "\n")
cat("RMSE:", round(rmse, 2), "\n")
cat("R2:", round(r2, 3), "\n")
```


1i) Response to part b. 

```{r}



```


## Question 2

2a) Response to part a. 


```{r}

# code for 2a

```


2b) Response to part b. 


```{r}

# code for 2b

```


## Question 3

Here is only text for question 3. 

If you want to add a hypertext link [you can do so](http://www.chapman.edu)

- If you want to add bullets
  - You
  - Can 
  - Do 
  - It

If you want to add an image [follow this guide](https://stackoverflow.com/questions/25166624/insert-picture-table-in-r-markdown)



